---
title: "part2"
author: "Lulu"
date: "7/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Descriptions

1. Load the ToothGrowth data and perform some basic exploratory data analyses
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by 
supp and dose.
4. State your conclusions and the assumptions needed for your conclusions.

## Point 1
1. Provide a basic summary of the data.

```{r, echo=FALSE}
str(ToothGrowth)
```

ToothGrowth is a "data.frame" containing 60 observations of 3 variables: "len", 
numeric; "supp", factor; "dose", numeric. 
There are two supplements: VC and OJ. Each one contains three doses at 0.5, 
1.0, and 2.0. There are 10 observations for each supp at each dose. 

```{r assign data, echo=FALSE}
# Assign ToothGrowth to a new variable tg
tg <- ToothGrowth
# Convert ToothGrowth$supp to character instead of factor
tg$supp <- as.character(tg$supp)
# Assign all VC and OJ to vc and oj
vc <- subset(tg,supp == "VC")
oj <- subset(tg, supp == "OJ")
```

```{r statistics, echo=FALSE}
# Load library
library(dplyr)
# Calculate means
mns <- tg%>%group_by(supp, dose)%>%summarise(mean(len))
mns
sds <- tg%>%group_by(supp, dose)%>%summarise(sd(len))
sds
```

As shown in the table, higher doses of VC or OJ induced more tooth growth. 
Maximum average of tooth growth 26.14 was observed at VC dose of 2.0. 

```{r plots, echo=FALSE}
# Make a plot of one row and 2 columns
par(mfrow=c(1,2), mar=c(4,4,2,1))

# Make boxplots of len for different doses of VC and OJ
boxplot(len~dose, vc, xlab="dose", ylab="ToothGrowth", main="VC", ylim=c(2,35))
boxplot(len~dose, oj, xlab="dose", ylab="ToothGrowth", main="OJ", ylim=c(2,35))
```

## Point 3
Use confidence intervals and/or hypothesis tests to compare tooth growth by supp
and dose.

```{r set up parameters, echo=FALSE}
vc_half <- subset(vc, dose==0.5)
vc_1 <- subset(vc, dose==1)
vc_2 <- subset(vc, dose==2)
oj_half <- subset(oj, dose==0.5)
oj_1 <- subset(oj, dose==1)
oj_2 <- subset(oj, dose==2)
```
1. Compare tooth growth associated with different doses for same supplement. 

1.1 Run t test to compare vc_half and vc_1. 
Hypothesis H0: same tooth growth at vc_1 and vc_half; Ha: more tooth growth at 
vc_1 compared to vc_half.
Assumes same variance in two groups.

```{r}
t.test(vc_1$len-vc_half$len, alternative="greater")
```

One sample t-test results show the 95 percent confidence interval of 
vc_1$len-vc_half$len is 6.16418 to Inf. Therefore, more tooth growth is 
associated with vc_1 compared to vc_half. 

1.2 Run t test to compare vc_2 and vc_1. 
Hypothesis H0: same tooth growth at vc_2 and vc_1; Ha: more tooth growth at 
vc_2 compared to vc_1.
Assumes same variance in two groups.

```{r}
t.test(vc_2$len-vc_1$len, alternative="greater")
```

One sample t-test results show the 95 percent confidence interval of 
vc_2$len-vc_1$len is 6.157075 to Inf. Therefore, more tooth growth is 
associated with vc_2 compared to vc_1. 

1.3 Run t test to compare oj_half and oj_1. 
Hypothesis H0: same tooth growth at oj_1 and oj_half; Ha: more tooth growth 
at oj_1 compared to oj_half.
Assumes same variance in two groups.

```{r}
t.test(oj_1$len-oj_half$len, alternative="greater")
```

One sample t-test results show the 95 percent confidence interval of 
oj_1$len-oj_half$len is 5.300497 to Inf. Therefore, more tooth growth is 
associated with oj_1 compared to oj_half. 

1.4 Run t test to compare oj_2 and oj_1. 
Hypothesis H0: same tooth growth at oj_2 and oj_1; Ha: more tooth growth at 
oj_2 compared to oj_1.
Assumes same variance in two groups.

```{r}
t.test(oj_2$len-oj_1$len, alternative="greater")
```

One sample t-test results show the 95 percent confidence interval of 
oj_2$len-oj_1$len is 0.1908169 to Inf. Therefore, more tooth growth is 
associated with oj_2 compared to oj_1. 

Conclusion: more tooth growth is associated with higher doses for the same 
supplement. 

2. Compare tooth growth associated with different supplement at the same doses. 

2.1 Run t test to compare vc_half and oj_half. 
Hypothesis H0: same tooth growth at vc_half and oj_half; Ha: more tooth growth 
at oj_half compared to vc_half.
Assumes same variance in two groups.

```{r}
t.test(oj_half$len-vc_half$len, alternative="greater")
```

One sample t-test results show the 95 percent confidence interval of 
oj_half$len-vc_half$len is 2.019552 to Inf. Therefore, more tooth growth is 
ssociated with oj_half compared to vc_half. 

2.2 Run t test to compare vc_1 and oj_1. 
Hypothesis H0: same tooth growth at vc_1 and oj_1; Ha: more tooth growth at 
oj_1 compared to vc_1.
Assumes same variance in two groups.

```{r}
t.test(oj_1$len-vc_1$len, alternative="greater")
```

One sample t-test results show the 95 percent confidence interval of 
oj_1$len-vc_1$len is 2.706401 to Inf. Therefore, more tooth growth is 
associated with oj_1 compared to vc_1. 

3.3 Run t test to compare vc_2 and oj_2. 
Hypothesis H0: same tooth growth at vc_2 and oj_2; Ha: different tooth growth 
at vc_2 compared to oj_2.
Assumes same variance in two groups.

```{r}
# Assumes same variance in two groups
t.test(vc_2$len-oj_2$len, alternative="two.sided")
```

One sample t-test results show the 95 percent confidence interval of 
vc_2$len-oj_2$len is -4.168976 to 4.328976. Therefore, accepting null hopothesis 
that same tooth growth at vc_2 and oj_2. 
